Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-07-14 prop_letter.tex
prop_letter.tex Small things here and there...

Contrary to the case with conventional quantum probability spaces,
there is generally no unique state consistent with a QIVPM. Given
observations done with finite resources, the probabilities become
intervals, and we say a state\nb$\rho$ is consistent with the
measure\nb$\bar{\mu}$ acting on a projector\nb$P$ if
$\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)$,
where $\mu_{\rho}^{\mathrm{B}}\left(P\right)=\Tr\left(\rho P\right)$
is the probability that the event\nb$P$ happens determined by the
Born rule\nb\citep{Born1983,peres1995quantum,544199}. Given a set
of projectors\nb$S$, 
\begin{equation}
\coreBorn\left(\bar{\mu},S\right)=\set{\textrm{mixed state }\rho}{\forall P\in S\ \mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)}
\end{equation}
is the set of all mixed states consistent with $\bar{\mu}$ on every
projector in $S$. For example, assuming the three intervals $\imposs=\left[0,0\right]$,
$\necess=\left[1,1\right]$, and $\unknown=\left[0,1\right]$, then:
\begin{equation}
\bar{\mu}_{1}\left(P\right)=\begin{cases}
\necess\,, & \text{if }P=\mathbb{1}\text{ ;}\\
\imposs\,, & \text{if }P=\mathbb{0}\text{ ;}\\
\unknown\,, & \text{otherwise,}
\end{cases}\label{eq:no-information-IVPM}
\end{equation}
is a QIVPM. This measure represents the beliefs of an experimenter
with no prior knowledge about the particular quantum system in question;
any quantum state would be consistent with $\bar{\mu}_{1}$ on any
projectors, and $\coreBorn\left(\bar{\mu}_{1},\events\right)$ is
just the set of all mixed states.

@@

Our first result is that a QIVPM is consistent with the inability to avoid
uncertainty found in conventional infinite precise quantum
mechanics\nb\cite{PeresRon1988,peres1995quantum}. We will adapt
the proof of the Kochen-Specker theorem~\cite{kochenspecker1967,
  peres1995quantum,Redhead1987-REDINA,Griffiths2003} to show that it
is, in general, impossible to have just the two intervals \imposs\ and
\necess: every QIVPM must be parameterized by a collection of
intervals that includes at least one additional ``fuzzy'' interval and
some events must be associated with such uncertain intervals. The
significance of this observation is that the uncertainty in the
intervals is, in contrast to the classical case, \emph{a fundamental
  part of quantum theory}. There are in fact two sources of
uncertainty: one due to the finite measurement resources and one due
to complementarity, which requires every quantum framework to be truly
probabilistic.

@@

\begin{example}\label{ex:three-dimensional-three-value}
  Given a three-dimensional Hilbert space with an orthonormal basis
  $\left\{ \ket{0},\ket{1},\ket{2}\right\} $,  let
  $\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
  $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
  $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}_{3}(\mathbb{0})=\bar{\mu}_{3}(\proj{0})\\
 & =\bar{\mu}_{3}(\proj{\ps})=\bar{\mu}_{3}(\proj{\ps'})\,,
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}_{3}(\mathbb{1})=\bar{\mu}_{3}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}_{3}(\mathbb{1}-\proj{\ps})=\bar{\mu}_{3}(\mathbb{1}-\proj{\ps'})\,,
\end{aligned}
\\
 &  & \unknown=\bar{\mu}_{3}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}
It is straightforward to verify that $\bar{\mu}_{3}$ is a QIVPM. Then we will
prove $\coreBorn\left(\bar{\mu}_{3},\events\right)$ is an empty set by
contradiction. Suppose there is a mixed
state\nb$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu}_{3},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since $\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}_{3}(P)$
for any\nb$P$, the mixed state\nb$\rho$ must satisfy
\begin{equation}
\melem{0}{\rho}{0}=\melem{\ps}{\rho}{\ps}=\melem{\ps'}{\rho}{\ps'}=0
\end{equation}
which implies $\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$
for all $j$. However, the last equation never holds. 
Thus, there is no state consistent with the QIVPM on all 
projectors\nb$P\in\events$, so Gleason's theorem does not hold for at least this
example.
\end{example}

>>>>

Contrary to the case with conventional quantum probability spaces,
there is generally no unique state consistent with a QIVPM. Given
observations done with finite resources, the probabilities become
intervals, and we say a state\nb$\rho$ is consistent with the
measure\nb$\bar{\mu}$ acting on a projector\nb$P$ if
$\mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)$,
where $\mu_{\rho}^{\mathrm{B}}\left(P\right)=\Tr\left(\rho P\right)$
is the probability that the event\nb$P$ happens determined by the
Born rule\nb\citep{Born1983,peres1995quantum,544199}. Given a set
of projectors\nb$S$, 
\begin{equation}
\coreBorn\left(\bar{\mu},S\right)=\set{\textrm{mixed state }\rho}{\forall P\in S\ \mu_{\rho}^{\mathrm{B}}\left(P\right)\in\bar{\mu}\left(P\right)}
\end{equation}
is the set of all mixed states consistent with $\bar{\mu}$ on every
projector in $S$. For example, assuming the three intervals $\imposs=\left[0,0\right]$,
$\necess=\left[1,1\right]$, and $\unknown=\left[0,1\right]$, then:
\begin{equation}
\bar{\mu}_{1}\left(P\right)=\begin{cases}
\imposs\,, & \text{if }P=\mathbb{0}\text{ ;}\\
\necess\,, & \text{if }P=\mathbb{1}\text{ ;}\\
\unknown\,, & \text{otherwise,}
\end{cases}\label{eq:no-information-IVPM}
\end{equation}
is a QIVPM. This measure represents the beliefs of an experimenter
with no prior knowledge about the particular quantum system in question;
any quantum state would be consistent with $\bar{\mu}_{1}$ on any
projectors, and $\coreBorn\left(\bar{\mu}_{1},\events\right)$ is
just the set of all mixed states.

@@

Our first result is that a QIVPM is consistent with the inability to avoid
uncertainty found in conventional infinite precise quantum
mechanics\nb\cite{PeresRon1988,peres1995quantum}. We will adapt
the proof of the Kochen-Specker theorem\nb\cite{kochenspecker1967,
  peres1995quantum,Redhead1987-REDINA,Griffiths2003} to show that it
is, in general, impossible to have just the two intervals \imposs\ and
\necess: every QIVPM must be parameterized by a collection of
intervals that includes at least one additional ``fuzzy'' interval and
some events must be associated with such uncertain intervals. The
significance of this observation is that the uncertainty in the
intervals is, in contrast to the classical case, \emph{a fundamental
  part of quantum theory}. There are in fact two sources of
uncertainty: one due to the finite measurement resources and one due
to complementarity, which requires every quantum framework to be truly
probabilistic.

@@

\begin{example}\label{ex:three-dimensional-three-value}
  Given a three-dimensional Hilbert space with an orthonormal basis
  $\left\{ \ket{0},\ket{1},\ket{2}\right\} $,  let
  $\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
  $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
  $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\,,
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\,,
\end{aligned}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}
It is straightforward to verify that $\bar{\mu}$ is a QIVPM. Then we will
prove $\coreBorn\left(\bar{\mu},\events\right)$ is an empty set by
contradiction. Suppose there is a mixed
state\nb$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since $\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}(P)$
for any\nb$P$, the mixed state\nb$\rho$ must satisfy
\begin{equation}
\melem{0}{\rho}{0}=\melem{\ps}{\rho}{\ps}=\melem{\ps'}{\rho}{\ps'}=0
\end{equation}
which implies $\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$
for all $j$. However, the last equation never holds. 
Thus, there is no state consistent with the QIVPM on all 
projectors\nb$P\in\events$, so Gleason's theorem does not hold for at least this
example.
\end{example}
